---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(unsum)
```

## (...)

## Caveat

It is strongly discouraged to run custom data transformations on the output of `closure_combine()` before passing it into other `closure_*()` functions. The latter need input with a very specific format, and if you manipulate the data between two `closure_*()` calls, these assumptions may no longer hold. Some checks are in place to detect alterations, but they may not catch all of them.

If you want to run your own transformations or analyses on `closure_combine()` output, it might be best to first save this output in a new variable specifically for use in your custom code. This way, you can still safely run unsum functions like `closure_plot_bar()` on the original data.
