---
title: "sprite"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sprite}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r include=FALSE}
devtools::load_all()
```

# Comparison to SPRITE (rsprite2)

```{r setup}
library(closure)
library(rsprite2)
library(purrr)
```

**TODO: Figure out why rsprite2 finds more distributions than closure!
Did I specify it incorrectly?**

## SPRITE

SPRITE finds 12 solutions. (The code asks for 20, so 12 is all it can
find.)

```{r}
solution_sprite <- set_parameters(
  mean = 3.5,
  sd = 0.5,
  n_obs = 52,
  min_val = 1,
  max_val = 5,
  m_prec = 1,
  sd_prec = 1
) %>% 
  find_possible_distributions(
    n_distributions = 20,
    seed = 1234
  )

solution_sprite

solution_sprite %>% 
  plot_distributions()
```

## CLOSURE

```{r}
solution_closure <- closure_combine(
  mean = "3.5",
  sd = "0.5",
  n = 52,
  scale_min = 1,
  scale_max = 5 
)

solution_closure

solution_closure %>% 
  closure_plot_bar()
```

## Comparison

SPRITE finds 12 distributions, CLOSURE only finds 3:

```{r}

nrow(solution_sprite)

nrow(solution_closure$results)
```

All solutions found by CLOSURE are also found by SPRITE:

```{r}
combos_sprite <- solution_sprite$distribution %>% 
  map(sort)

combos_closure <- solution_closure$results$combination %>% 
  map(sort)

closure_combos_are_subset <- combos_closure %in% combos_sprite
closure_combos_are_subset

all(closure_combos_are_subset)
```

However, if `m_prec` and `sd_prec` are `2` in rsprite2's
`set_parameters()`, and `mean` and `sd` in `closure_combine()` have two
decimal places each, both approaches only find a single distribution —
and it is the same one CLOSURE finds! (The SPRITE and CLOSURE
distributions happen to be identical even without `sort()`, but I
wouldn't rely on this.)

The trouble is that the two sets of combinations are not the same with
only one decimal place. If SPRITE can find them, shouldn't CLOSURE find
them, too? Or am I missing something about the way I specified the
rsprite2 functions?
