---
title: "malbouby2025"
output: html_document
---

```{r include=FALSE}
devtools::load_all(".")
```

[Malbouby et al. (2025)](https://nutrition.bmj.com/content/early/2025/02/06/bmjnph-2024-000997) reported red flags in a paper published by a BMJ journal. The publication process of this comment inspired James Heathers, one of its authors, to pen a remarkable testimony ([Heathers 2025](https://jamesclaims.substack.com/p/research-integrity-is-a-clown-car?r=6xosj&utm_campaign=post&utm_medium=web&triedRedirect=true)).

Among the concerns raised by Malbouby et al. (2025) was the shape of the sample reconstructed from summary statistics in the original paper using the SPRITE technique. The average reconstructed sample showed a large amount of variation approximating a bimodal distribution, and indeed what [Heathers (2017)](https://jamesheathers.medium.com/sprite-case-study-3-soup-is-good-albeit-extremely-confusing-food-96ea526c488d) referred to as the "horns of no confidence": a much larger degree of polarization towards the scale limits than what would be expected on theoretical grounds (in this case, human body mass index).

Here, I want to use CLOSURE, a complete generalization of the probabilistic SPRITE algorithm, to assess the SPRITE analysis by Malbouby et al. (2025). Whereas SPRITE only creates a random assortment of samples that could have lead to the reported summary statistics, CLOSURE always finds all of those possible samples.

```{r}
result <- closure_generate(
  mean = "30.6",
  sd = "3.1",
  n = 30,
  scale_min = 27,
  scale_max = 34
)
```

The results in detail:

```{r}
result
```

Among the metrics, a particularly interesting quantity is the horns index, $h$. It ranges from 0 to 1, where 0 means that all observations have the same value, and 1 represents a sample that is evenly split between the scale limits, with no observations in between. In this way, $h$ quantifies the variability among the reconstructed values given their scale limits. Here, $h$ is 0.76, a relatively high value. For reference, a uniform distribution with 8 scale values, such as those from 27 to 34 here, would only have an $h$ of 0.43, as the `horns_uniform` column shows.

This is remarkable given that one would not expect a uniform distribution, let alone a sample as polarized as the one implied by the summary statistics, to result from a BMI measurement. Rather, one would expect a stronger central tendency.

Visualizing the results:

```{r}
closure_plot_bar(result, frequency = "percent")
closure_plot_ecdf(result)
```
