---
title: "Exploring boredom bias in SPRITE"
output:
  html_document: default
  pdf_document: default
---

# Summary

Here, I investigate a possible kind of bias in SPRITE. When comparing sets of samples generated by SPRITE to the respective complete set generated by CLOSURE, it seems that SPRITE is biased towards "boring" results. These have at least one of the following characteristics:

-   They tend to have lower variance than a complete set of results.

-   The difference between their highest and lowest normalized variance (horns index) is lower.

-   In cases where the histogram of the complete results set found by CLOSURE has a slightly V-shaped indentation about at the scale midpoint, the difference in frequency between any two adjacent non-extreme values in the respective SPRITE plot tends to be incremental, and to form a staircase-like pattern in a roughly monotonous fashion. This does not necessarily mean that CLOSURE finds higher variance, as the low-variance end in test case 2 shows. Here, SPRITE shows the staircase pattern but actually has higher variance than CLOSURE, which shows the V-shape. Rather, this phenomenon is a genuinely distinct issue above and beyond the low-variance bias noted above: a local boredom pattern at the non-extreme values.

Claude offered a plausible explanation of these observed phenomena (skip to Cause 2; the first one is irrelevant and now fixed): <https://github.com/lhdjung/closure-core/blob/master/protocols/sprite-bias-assessment.md>

# Analysis

The development version is currently needed to run this R Markdown file, but it might fail. I only just noticed and I'm not sure why.

```{r}
# remotes::install_github("lhdjung/unsum")
library(unsum)
library(ggplot2)
```

Generate SPRITE and CLOSURE example data first. Detailed results are omitted here.

```{r message=FALSE}
data1_sprite <- sprite_generate(mean = "3.0", sd = "1.0", n = 100, scale_min = 1, scale_max = 5)
data2_sprite <- sprite_generate(mean = "4.7", sd = "2.2", n = 50, scale_min = 1, scale_max = 7)
data3_sprite <- sprite_generate(mean = "2.00", sd = "1.45", n = 65, scale_min = 1, scale_max = 5)
data4_sprite <- sprite_generate(mean = "4.63", sd = "1.08", n = 72, scale_min = 1, scale_max = 7)
data5_sprite <- sprite_generate(mean = "2.16", sd = "1.02", n = 114, scale_min = 1, scale_max = 5)

data1_closure <- closure_generate(mean = "3.0", sd = "1.0", n = 100, scale_min = 1, scale_max = 5)
data2_closure <- closure_generate(mean = "4.7", sd = "2.2", n = 50, scale_min = 1, scale_max = 7)
data3_closure <- closure_generate(mean = "2.00", sd = "1.45", n = 65, scale_min = 1, scale_max = 5)
data4_closure <- closure_generate(mean = "4.63", sd = "1.08", n = 72, scale_min = 1, scale_max = 7)
data5_closure <- closure_generate(mean = "2.16", sd = "1.02", n = 114, scale_min = 1, scale_max = 5)
```

## Count the samples found by SPRITE and CLOSURE

### Test case 1

```{r}
data1_sprite$metrics_main$samples_all
data1_closure$metrics_main$samples_all
```

### Test case 2

```{r}
data2_sprite$metrics_main$samples_all
data2_closure$metrics_main$samples_all
```

### Test case 3

```{r}
data3_sprite$metrics_main$samples_all
data3_closure$metrics_main$samples_all
```

### Test case 4

```{r}
data4_sprite$metrics_main$samples_all
data4_closure$metrics_main$samples_all
```

### Test case 5

```{r}
data5_sprite$metrics_main$samples_all
data5_closure$metrics_main$samples_all
```

## Visualize the average min-max variance samples

### Test case 1

```{r}
sprite_plot_bar(data1_sprite) + labs(title = "SPRITE test case 1")
closure_plot_bar(data1_closure) + labs(title = "CLOSURE test case 1")
```

### Test case 2

```{r}
sprite_plot_bar(data2_sprite) + labs(title = "SPRITE test case 2")
closure_plot_bar(data2_closure) + labs(title = "CLOSURE test case 2")
```

### Test case 3

```{r}
sprite_plot_bar(data3_sprite) + labs(title = "SPRITE test case 3")
closure_plot_bar(data3_closure) + labs(title = "CLOSURE test case 3")
```

### Test case 4

```{r}
sprite_plot_bar(data4_sprite) + labs(title = "SPRITE test case 4")
closure_plot_bar(data4_closure) + labs(title = "CLOSURE test case 4")
```

### Test case 5

```{r}
sprite_plot_bar(data5_sprite) + labs(title = "SPRITE test case 5")
closure_plot_bar(data5_closure) + labs(title = "CLOSURE test case 5")
```
